ABCD_OHBM_PFN_Analyses

### 2022-02-21 ABCD_OHBM_PFN_Analyses.Rmd
random effects
included random effects for both family and site in ABCD analyses

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(width=250)
library(tidyr)
library(dplyr)
library(ggplot2)
library(lsr)
library(lme4)
library(lmerTest)
```

@@

combined.data <- merge(data.clean,abcd_imgincl01[, c("subjectkey",'imgincl_t1w_include')],by="subjectkey")
```

@@

# Remove outliers
mean_PC1 <- mean(combined.data$thompson_PC1,na.rm=TRUE)
mean_PC2 <- mean(combined.data$thompson_PC2,na.rm=TRUE)
mean_PC3 <- mean(combined.data$thompson_PC3,na.rm=TRUE)
sd3_PC1 <- sd(combined.data$thompson_PC1,na.rm=TRUE)*3
sd3_PC2 <- sd(combined.data$thompson_PC2,na.rm=TRUE)*3
sd3_PC3 <- sd(combined.data$thompson_PC3,na.rm=TRUE)*3
sum(combined.data$thompson_PC1 > mean_PC1+sd3_PC1,na.rm=TRUE) # 8 outliers
sum(combined.data$thompson_PC2 > mean_PC2+sd3_PC2,na.rm=TRUE) # 2 outliers
sum(combined.data$thompson_PC3 > mean_PC3+sd3_PC3,na.rm=TRUE) # 0 outliers
combined.data.noOutliers <- combined.data[combined.data$thompson_PC1>(mean_PC1-sd3_PC1) & combined.data$thompson_PC1<(mean_PC1+sd3_PC1) & combined.data$thompson_PC2>(mean_PC2-sd3_PC2) & combined.data$thompson_PC2<(mean_PC2+sd3_PC2) & combined.data$thompson_PC3>(mean_PC3-sd3_PC3) & combined.data$thompson_PC3<(mean_PC3+sd3_PC3),]
```

@@

combined.data.noOutliers[,224:240]<-as.numeric(unlist(combined.data.noOutliers[,224:240]))

# Calculation of subscore with items related to physical and sexual abuse
combined.data.noOutliers$abuse <- rowSums(data.frame(as.numeric(combined.data.noOutliers$ksads_ptsd_raw_761_p),as.numeric(combined.data.noOutliers$ksads_ptsd_raw_762_p),as.numeric(combined.data.noOutliers$ksads_ptsd_raw_763_p),as.numeric(combined.data.noOutliers$ksads_ptsd_raw_767_p),as.numeric(combined.data.noOutliers$ksads_ptsd_raw_768_p),as.numeric(combined.data.noOutliers$ksads_ptsd_raw_769_p)),na.rm=TRUE)

# Calculation of subscore with items related to trauma (car accidents, natural disasters, loss, etc.)
combined.data.noOutliers$trauma <- rowSums(data.frame(as.numeric(combined.data.noOutliers$ksads_ptsd_raw_754_p),as.numeric(combined.data.noOutliers$ksads_ptsd_raw_755_p),as.numeric(combined.data.noOutliers$ksads_ptsd_raw_756_p),as.numeric(combined.data.noOutliers$ksads_ptsd_raw_757_p),as.numeric(combined.data.noOutliers$ksads_ptsd_raw_758_p),as.numeric(combined.data.noOutliers$ksads_ptsd_raw_759_p),as.numeric(combined.data.noOutliers$ksads_ptsd_raw_760_p),as.numeric(combined.data.noOutliers$ksads_ptsd_raw_766_p),as.numeric(combined.data.noOutliers$ksads_ptsd_raw_770_p)),na.rm=TRUE)

# Remember to add in this variable about family structure so we can use it as a covariate
family <-read.table("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/ses/acspsw03.txt",header=TRUE)
family.baseline<-family[family$eventname=="baseline_year_1_arm_1",]
abcd.data <- merge(combined.data.noOutliers,family.baseline[, c("subjectkey", setdiff(colnames(family.baseline),colnames(combined.data.noOutliers)))],by="subjectkey")

emo.neglect <-read.table("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/neglect/crpbi01.txt",header=TRUE)
emo.neglect.baseline<-emo.neglect[emo.neglect$eventname=="baseline_year_1_arm_1",]
phy.neglect <-read.table("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/neglect/pmq01.txt",header=TRUE)
phy.neglect.baseline<-phy.neglect[phy.neglect$eventname=="baseline_year_1_arm_1",]
neglect.data <- merge(emo.neglect.baseline,phy.neglect.baseline[, c("subjectkey", setdiff(colnames(phy.neglect.baseline),colnames(emo.neglect.baseline)))],by="subjectkey")

abcd.data<-merge(abcd.data,neglect.data[,c("subjectkey", setdiff(colnames(neglect.data),colnames(abcd.data)))],by="subjectkey")

# Calculate subscales of deprivation (emotional and physical neglect) - see this ref for motivation https://www.sciencedirect.com/science/article/pii/S2352289518300821
abcd.data$acceptance <- rowSums(data.frame(as.numeric(abcd.data$crpbi_parent1_y),as.numeric(abcd.data$crpbi_parent2_y),as.numeric(abcd.data$crpbi_parent3_y),as.numeric(abcd.data$crpbi_parent4_y),as.numeric(abcd.data$crpbi_parent5_y)))

abcd.data$monitoring <- rowSums(data.frame(as.numeric(abcd.data$parent_monitor_q1_y),as.numeric(abcd.data$parent_monitor_q2_y),as.numeric(abcd.data$parent_monitor_q3_y),as.numeric(abcd.data$parent_monitor_q4_y),as.numeric(abcd.data$parent_monitor_q5_y)))

abcd.data$abuse.binary <- factor(abcd.data$abuse>=1,labels = c("No Abuse","Abuse"))
abcd.data$abuse.binary.numeric <- as.numeric(abcd.data$abuse.binary)
abcd.data$trauma.binary <- factor(abcd.data$trauma>=1,labels = c("No Trauma","Trauma"))
abcd.data$trauma.binary.numeric <- as.numeric(abcd.data$trauma.binary)

median.monitoring = median(na.omit(abcd.data$monitoring))
median.acceptance = median(na.omit(abcd.data$acceptance))
abcd.data$monitoring.binary <- factor(abcd.data$monitoring<20,labels = c("Low Physical Neglect","High Physical Neglect"))
abcd.data$monitoring.binary.numeric <- as.numeric(abcd.data$monitoring.binary)
abcd.data$acceptance.binary <- factor(abcd.data$acceptance<7.5,labels = c("Low Emotional Neglect","High Emotional Neglect"))
abcd.data$acceptance.binary.numeric <- as.numeric(abcd.data$acceptance.binary)

# calculate a total FPN score
combined.data.noOutliers$totalFPN <- rowSums(data.frame(as.numeric(combined.data.noOutliers$df4),as.numeric(combined.data.noOutliers$df16),as.numeric(combined.data.noOutliers$df18)))

# separate train and test sets
trainIDs <- read.csv("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/train.csv")
train.data <- merge(combined.data.noOutliers,trainIDs,by="subjectkey") # n=2,149
test.data <- combined.data.noOutliers[!is.element(combined.data.noOutliers$subjectkey,trainIDs$subjectkey),] #n=2,424
```

@@

@@

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(width=250)
library(tidyr)
library(dplyr)
library(ggplot2)
library(lsr)
library(lme4)
library(lmerTest)
library(mediation)
library(ggcorrplot)
library(corrplot)
library(RColorBrewer)
library(glmnet)
library(mgcv)
library(gamm4)
```

@@

combined.data <- merge(data.clean,abcd_imgincl01[, c("subjectkey",'imgincl_t1w_include')],by="subjectkey") # n=4,614

@@

# Remove outliers
mean_PC1 <- mean(combined.data$thompson_PC1,na.rm=TRUE)
mean_PC2 <- mean(combined.data$thompson_PC2,na.rm=TRUE)
mean_PC3 <- mean(combined.data$thompson_PC3,na.rm=TRUE)
sd3_PC1 <- sd(combined.data$thompson_PC1,na.rm=TRUE)*3
sd3_PC2 <- sd(combined.data$thompson_PC2,na.rm=TRUE)*3
sd3_PC3 <- sd(combined.data$thompson_PC3,na.rm=TRUE)*3
sum(combined.data$thompson_PC1 > mean_PC1+sd3_PC1,na.rm=TRUE) # 8 outliers
sum(combined.data$thompson_PC2 > mean_PC2+sd3_PC2,na.rm=TRUE) # 2 outliers
sum(combined.data$thompson_PC3 > mean_PC3+sd3_PC3,na.rm=TRUE) # 0 outliers
combined.data.noOutliers <- combined.data[combined.data$thompson_PC1>(mean_PC1-sd3_PC1) & combined.data$thompson_PC1<(mean_PC1+sd3_PC1) & combined.data$thompson_PC2>(mean_PC2-sd3_PC2) & combined.data$thompson_PC2<(mean_PC2+sd3_PC2) & combined.data$thompson_PC3>(mean_PC3-sd3_PC3) & combined.data$thompson_PC3<(mean_PC3+sd3_PC3),]

# n=4,579
```

@@

# combined.data.noOutliers[,224:240]<-as.numeric(unlist(combined.data.noOutliers[,224:240]))
# 
# # Calculation of subscore with items related to physical and sexual abuse
# combined.data.noOutliers$abuse <- rowSums(data.frame(as.numeric(combined.data.noOutliers$ksads_ptsd_raw_761_p),as.numeric(combined.data.noOutliers$ksads_ptsd_raw_762_p),as.numeric(combined.data.noOutliers$ksads_ptsd_raw_763_p),as.numeric(combined.data.noOutliers$ksads_ptsd_raw_767_p),as.numeric(combined.data.noOutliers$ksads_ptsd_raw_768_p),as.numeric(combined.data.noOutliers$ksads_ptsd_raw_769_p)),na.rm=TRUE)
# 
# # Calculation of subscore with items related to trauma (car accidents, natural disasters, loss, etc.)
# combined.data.noOutliers$trauma <- rowSums(data.frame(as.numeric(combined.data.noOutliers$ksads_ptsd_raw_754_p),as.numeric(combined.data.noOutliers$ksads_ptsd_raw_755_p),as.numeric(combined.data.noOutliers$ksads_ptsd_raw_756_p),as.numeric(combined.data.noOutliers$ksads_ptsd_raw_757_p),as.numeric(combined.data.noOutliers$ksads_ptsd_raw_758_p),as.numeric(combined.data.noOutliers$ksads_ptsd_raw_759_p),as.numeric(combined.data.noOutliers$ksads_ptsd_raw_760_p),as.numeric(combined.data.noOutliers$ksads_ptsd_raw_766_p),as.numeric(combined.data.noOutliers$ksads_ptsd_raw_770_p)),na.rm=TRUE)
# 
# # Remember to add in this variable about family structure so we can use it as a covariate
# family <-read.table("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/ses/acspsw03.txt",header=TRUE)
# family.baseline<-family[family$eventname=="baseline_year_1_arm_1",]
# abcd.data <- merge(combined.data.noOutliers,family.baseline[, c("subjectkey", setdiff(colnames(family.baseline),colnames(combined.data.noOutliers)))],by="subjectkey")
# 
# emo.neglect <-read.table("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/neglect/crpbi01.txt",header=TRUE)
# emo.neglect.baseline<-emo.neglect[emo.neglect$eventname=="baseline_year_1_arm_1",]
# phy.neglect <-read.table("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/neglect/pmq01.txt",header=TRUE)
# phy.neglect.baseline<-phy.neglect[phy.neglect$eventname=="baseline_year_1_arm_1",]
# neglect.data <- merge(emo.neglect.baseline,phy.neglect.baseline[, c("subjectkey", setdiff(colnames(phy.neglect.baseline),colnames(emo.neglect.baseline)))],by="subjectkey")
# 
# abcd.data<-merge(abcd.data,neglect.data[,c("subjectkey", setdiff(colnames(neglect.data),colnames(abcd.data)))],by="subjectkey")
# 
# # Calculate subscales of deprivation (emotional and physical neglect) - see this ref for motivation https://www.sciencedirect.com/science/article/pii/S2352289518300821
# abcd.data$acceptance <- rowSums(data.frame(as.numeric(abcd.data$crpbi_parent1_y),as.numeric(abcd.data$crpbi_parent2_y),as.numeric(abcd.data$crpbi_parent3_y),as.numeric(abcd.data$crpbi_parent4_y),as.numeric(abcd.data$crpbi_parent5_y)))
# 
# abcd.data$monitoring <- rowSums(data.frame(as.numeric(abcd.data$parent_monitor_q1_y),as.numeric(abcd.data$parent_monitor_q2_y),as.numeric(abcd.data$parent_monitor_q3_y),as.numeric(abcd.data$parent_monitor_q4_y),as.numeric(abcd.data$parent_monitor_q5_y)))
# 
# abcd.data$abuse.binary <- factor(abcd.data$abuse>=1,labels = c("No Abuse","Abuse"))
# abcd.data$abuse.binary.numeric <- as.numeric(abcd.data$abuse.binary)
# abcd.data$trauma.binary <- factor(abcd.data$trauma>=1,labels = c("No Trauma","Trauma"))
# abcd.data$trauma.binary.numeric <- as.numeric(abcd.data$trauma.binary)
# 
# median.monitoring = median(na.omit(abcd.data$monitoring))
# median.acceptance = median(na.omit(abcd.data$acceptance))
# abcd.data$monitoring.binary <- factor(abcd.data$monitoring<20,labels = c("Low Physical Neglect","High Physical Neglect"))
# abcd.data$monitoring.binary.numeric <- as.numeric(abcd.data$monitoring.binary)
# abcd.data$acceptance.binary <- factor(abcd.data$acceptance<7.5,labels = c("Low Emotional Neglect","High Emotional Neglect"))
# abcd.data$acceptance.binary.numeric <- as.numeric(abcd.data$acceptance.binary)
```

# More data prep (family ID, site ID, total FPN score, train/test)
```{r}
# calculate a total FPN score
combined.data$totalFPN <- rowSums(data.frame(as.numeric(combined.data$df4),as.numeric(combined.data$df16),as.numeric(combined.data$df18)))

# Remember to add in this variable about family structure so we can use it as a covariate
family <-read.table("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/ses/acspsw03.txt",header=TRUE)
family.baseline<-family[family$eventname=="baseline_year_1_arm_1",]
abcd.data.almost <- merge(combined.data,family.baseline[, c("subjectkey", setdiff(colnames(family.baseline),colnames(combined.data)))],by="subjectkey")

# also add in the variable for site ID to use as a covariate
site_info <- readRDS("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/DEAP-siteID.rds")
site.baseline<-site_info[site_info$event_name=="baseline_year_1_arm_1",]
colnames(site.baseline)[1]<-"subjectkey"
abcd.data.withFamAndSite <- merge(abcd.data.almost,site.baseline[,c("subjectkey", setdiff(colnames(site.baseline),colnames(abcd.data.almost)))],by="subjectkey")

# separate train and test sets
trainIDs <- read.csv("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/train.csv")
train.data <- merge(abcd.data.withFamAndSite,trainIDs,by="subjectkey") # n=2,290
test.data <- abcd.data.withFamAndSite[!is.element(abcd.data.withFamAndSite$subjectkey,trainIDs$subjectkey),] #n=2,323
```

@@

# Network 17 is specifically associated with PC1 and not PC2 or PC3
summary(lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+(1|abcd_site)+df18,data=train.data)) #***
summary(lmer(thompson_PC2~interview_age+sex+(1|rel_family_id)+(1|abcd_site)+df18,data=train.data))
summary(lmer(thompson_PC3~interview_age+sex+(1|rel_family_id)+(1|abcd_site)+df18,data=train.data))

# Network 15 is specifically associated with PC1 and PC3 but not PC2
summary(lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+(1|abcd_site)+df16,data=train.data)) #**
summary(lmer(thompson_PC2~interview_age+sex+(1|rel_family_id)+(1|abcd_site)+df16,data=train.data))
summary(lmer(thompson_PC3~interview_age+sex+(1|rel_family_id)+(1|abcd_site)+df16,data=train.data)) #**

# Network 3 is not associated with any of the PCs
summary(lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+(1|abcd_site)+df4,data=train.data)) 
summary(lmer(thompson_PC2~interview_age+sex+(1|rel_family_id)+(1|abcd_site)+df4,data=train.data))
summary(lmer(thompson_PC3~interview_age+sex+(1|rel_family_id)+(1|abcd_site)+df4,data=train.data))
```

```{r}
# # We already know that trauma and physical neglect are associated with impairment in PC1
# # Physical neglect is also associated with impairment in PC3
# summary(lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+trauma.binary,data=train.data)) #**
# summary(lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+monitoring.binary,data=train.data)) #***
# summary(lmer(thompson_PC3~interview_age+sex+(1|rel_family_id)+monitoring.binary,data=train.data)) #***
# 
# 
# # Are any subtypes of ACEs associated with PFN topography?
# summary(lmer(totalFPN~interview_age+sex+(1|rel_family_id)+monitoring.binary,data=train.data)) # ns
# summary(lmer(totalFPN~interview_age+sex+(1|rel_family_id)+acceptance.binary,data=train.data)) # ns
# summary(lmer(totalFPN~interview_age+sex+(1|rel_family_id)+abuse.binary,data=train.data)) # ns
# summary(lmer(totalFPN~interview_age+sex+(1|rel_family_id)+trauma.binary,data=train.data)) # ns
# 
# summary(lmer(df18~interview_age+sex+(1|rel_family_id)+monitoring.binary,data=train.data)) # ns
# summary(lmer(df18~interview_age+sex+(1|rel_family_id)+acceptance.binary,data=train.data)) # ns
# summary(lmer(df18~interview_age+sex+(1|rel_family_id)+abuse.binary,data=train.data)) # ns
# summary(lmer(df18~interview_age+sex+(1|rel_family_id)+trauma.binary,data=train.data)) # ns
# 
# summary(lmer(df16~interview_age+sex+(1|rel_family_id)+monitoring.binary,data=train.data)) # ns (0.056)
# summary(lmer(df16~interview_age+sex+(1|rel_family_id)+acceptance.binary,data=train.data)) # ns
# summary(lmer(df16~interview_age+sex+(1|rel_family_id)+abuse.binary,data=train.data)) # ns
# summary(lmer(df16~interview_age+sex+(1|rel_family_id)+trauma.binary,data=train.data)) # ns
# 
# summary(lmer(df4~interview_age+sex+(1|rel_family_id)+monitoring.binary,data=train.data)) # ns
# summary(lmer(df4~interview_age+sex+(1|rel_family_id)+acceptance.binary,data=train.data)) # ns
# summary(lmer(df4~interview_age+sex+(1|rel_family_id)+abuse.binary,data=train.data)) # ns
# summary(lmer(df4~interview_age+sex+(1|rel_family_id)+trauma.binary,data=train.data)) # ns

@@

# Mediation analyses -- these are significant!
```{r}
# Mediation: trauma --> Network 17 --> PC1 (path C does decrease) *
summary(lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+trauma.binary,data=train.data))
summary(lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+trauma.binary+df18,data=train.data))

bs_indir = c()
for (i in 1:1000){
  thisSample <- sample(c(1:length(train.data$subjectkey)),length(train.data$subjectkey),replace=TRUE)
  pathc<-lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+trauma.binary,data=train.data[thisSample,])
  pathcp<-lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+trauma.binary+df18,data=train.data[thisSample,])
  indir<-fixef(pathc)[4]-fixef(pathcp)[5]
  bs_indir <- c(bs_indir,indir)
}
t.test(bs_indir,conf.level=0.95)

# Mediation: trauma --> Network 15 --> PC1 (path C does not decrease)
summary(lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+trauma.binary,data=train.data))
summary(lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+trauma.binary+df16,data=train.data))

# Mediation: physical neglect --> Network 17 --> PC1 (path C does decrease) *
summary(lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+monitoring.binary,data=train.data))
summary(lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+monitoring.binary+df18,data=train.data))

bs_indir = c()
for (i in 1:1000){
  thisSample <- sample(c(1:length(train.data$subjectkey)),length(train.data$subjectkey),replace=TRUE)
  pathc<-lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+monitoring.binary,data=train.data[thisSample,])
  pathcp<-lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+monitoring.binary+df18,data=train.data[thisSample,])
  indir<-fixef(pathc)[4]-fixef(pathcp)[5]
  bs_indir <- c(bs_indir,indir)
}
t.test(bs_indir,conf.level=0.95)

# Mediation: physical neglect --> Network 15 --> PC1 (path C does not decrease)
summary(lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+monitoring.binary,data=train.data))
summary(lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+monitoring.binary+df16,data=train.data))

# Mediation: physical neglect --> Network 17 --> PC3 (path C does decrease) *
summary(lmer(thompson_PC3~interview_age+sex+(1|rel_family_id)+monitoring.binary,data=train.data))
summary(lmer(thompson_PC3~interview_age+sex+(1|rel_family_id)+monitoring.binary+df18,data=train.data))

bs_indir = c()
for (i in 1:1000){
  thisSample <- sample(c(1:length(train.data$subjectkey)),length(train.data$subjectkey),replace=TRUE)
  pathc<-lmer(thompson_PC3~interview_age+sex+(1|rel_family_id)+monitoring.binary,data=train.data[thisSample,])
  pathcp<-lmer(thompson_PC3~interview_age+sex+(1|rel_family_id)+monitoring.binary+df18,data=train.data[thisSample,])
  indir<-fixef(pathc)[4]-fixef(pathcp)[5]
  bs_indir <- c(bs_indir,indir)
}
t.test(bs_indir,conf.level=0.95)

# Mediation: physical neglect --> Network 15 --> PC3 (path C does not decrease)
summary(lmer(thompson_PC3~interview_age+sex+(1|rel_family_id)+monitoring.binary,data=train.data))
summary(lmer(thompson_PC3~interview_age+sex+(1|rel_family_id)+monitoring.binary+df16,data=train.data))

# No moderations were significant
```{r}
# # Moderation: trauma * Network 17 --> PC1
# summary(lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+trauma.binary*df18,data=train.data))
# 
# # Moderation: trauma * Network 15 --> PC1
# summary(lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+trauma.binary*df16,data=train.data))
# 
# # Moderation: physical neglect * Network 17 --> PC1
# summary(lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+monitoring.binary*df18,data=train.data))
# 
# # Moderation: physical neglect * Network 15 --> PC1
# summary(lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+monitoring.binary*df16,data=train.data))
# 
# # Moderation: physical neglect * Network 17 --> PC3
# summary(lmer(thompson_PC3~interview_age+sex+(1|rel_family_id)+monitoring.binary*df18,data=train.data))
# 
# # Moderation: physical neglect * Network 15 --> PC3 
# summary(lmer(thompson_PC3~interview_age+sex+(1|rel_family_id)+monitoring.binary*df16,data=train.data))
# 
# 
# # Moderation: sex * trauma * Network 17 --> PC1
# summary(lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+trauma.binary*df18*sex,data=train.data))
# 
# # Moderation: sex * trauma * Network 15 --> PC1
# summary(lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+trauma.binary*df16*sex,data=train.data))
# 
# # Moderation: sex * physical neglect * Network 17 --> PC1
# summary(lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+monitoring.binary*df18*sex,data=train.data))
# 
# # Moderation: sex * physical neglect * Network 15 --> PC1
# summary(lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+monitoring.binary*df16*sex,data=train.data))
# 
# # Moderation: sex * physical neglect * Network 17 --> PC3
# summary(lmer(thompson_PC3~interview_age+sex+(1|rel_family_id)+monitoring.binary*df18*sex,data=train.data))
# 
# # Moderation: sex * physical neglect * Network 15 --> PC3 
# summary(lmer(thompson_PC3~interview_age+sex+(1|rel_family_id)+monitoring.binary*df16*sex,data=train.data))
```

# Analyses of total FPN
```{r}
# summary(lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+totalFPN,data=train.data)) #**
# summary(lmer(thompson_PC2~interview_age+sex+(1|rel_family_id)+totalFPN,data=train.data)) #ns
# summary(lmer(thompson_PC3~interview_age+sex+(1|rel_family_id)+totalFPN,data=train.data)) #ns
# 
# summary(lmer(totalFPN~interview_age+sex+(1|rel_family_id)+monitoring.binary,data=train.data)) #ns
# summary(lmer(totalFPN~interview_age+sex+(1|rel_family_id)+acceptance.binary,data=train.data)) #ns
# summary(lmer(totalFPN~interview_age+sex+(1|rel_family_id)+abuse.binary,data=train.data)) #ns
# summary(lmer(totalFPN~interview_age+sex+(1|rel_family_id)+trauma.binary,data=train.data)) #ns
# 
# summary(lmer(totalFPN~interview_age+sex+(1|rel_family_id)+monitoring.binary*sex,data=train.data)) #ns
# summary(lmer(totalFPN~interview_age+sex+(1|rel_family_id)+acceptance.binary*sex,data=train.data)) #ns
# summary(lmer(totalFPN~interview_age+sex+(1|rel_family_id)+abuse.binary*sex,data=train.data)) #ns
# summary(lmer(totalFPN~interview_age+sex+(1|rel_family_id)+trauma.binary*sex,data=train.data)) #ns

# Port data over to matlab (womp) so I can run the Canlab multilevel mediation toolbox 
```{r}
write.csv(train.data,file="trainData_forMatlab.csv")
write.csv(test.data,file="testData_forMatlab.csv")
```

# Create a correlation matrix of all the important variables
```{r}
# important.variables <- train.data[,c("interview_age","monitoring","acceptance","abuse","trauma","df2","df3","df4","df5","df6","df7","df8","df9","df10","df11","df12","df13","df14","df15","df16","df17","df18","thompson_PC1","thompson_PC2","thompson_PC3")]
# 
# colnames(important.variables)<-c("Age","Phys Neglect","Emo Neglect","Abuse","Trauma","PFN 1","PFN 2","PFN 3","PFN 4","PFN 5","PFN 6","PFN 7","PFN 8","PFN 9","PFN 10","PFN 11","PFN 12","PFN 13","PFN 14","PFN 15","PFN 16","PFN 17","General Cog (PC1)","Exec Fxn (PC2)","Learn/Mem (PC3)")
# 
# png("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/ABCD_PFNs_ACES_Cog_CorrMatrix.png",
#     width = 8000, 
#     height = 4000,
#     res=300)
# corrplot(cor(important.variables,use="complete.obs"),method=c("number"), tl.col = "black", col=brewer.pal(n=8, name="Spectral"))
# dev.off() 
```

# Create a correlation matrix of just the cognitive and PFN variables
```{r}
important.variables <- train.data[,c("interview_age","df2","df3","df4","df5","df6","df7","df8","df9","df10","df11","df12","df13","df14","df15","df16","df17","df18","thompson_PC1","thompson_PC2","thompson_PC3")]

colnames(important.variables)<-c("Age","PFN 1","PFN 2","PFN 3","PFN 4","PFN 5","PFN 6","PFN 7","PFN 8","PFN 9","PFN 10","PFN 11","PFN 12","PFN 13","PFN 14","PFN 15","PFN 16","PFN 17","General Cog (PC1)","Exec Fxn (PC2)","Learn/Mem (PC3)")

png("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/ABCD_PFNs_Cog_CorrMatrix.png",
    width = 8000, 
    height = 4000,
    res=300)
corrplot(cor(important.variables,use="complete.obs"),method=c("number"), tl.col = "black", col=brewer.pal(n=8, name="Spectral"))
dev.off() 
```

## Cui et al. 2020 replications
```{r}
correlation<-cor.test(combined.data[combined.data$df18>3000,]$thompson_PC1,combined.data[combined.data$df18>3000,]$df18)
png("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/CuiRep_PC1_PFN17.png",
    width = 4000, 
    height = 4000,
    res=1000)
ggplot(data=combined.data[combined.data$df18>3000,],aes(thompson_PC1,df18))+geom_point(size=1,colour='gold')+geom_smooth(method=lm,colour='gray')+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black")) + ylab('Network 17') + xlab('Thompson PC1: General Cognition') + ggtitle(paste("r=",as.character(round(correlation$estimate,3)),", p=",as.character(round(correlation$p.value,3))))
dev.off()

correlation <- cor.test(combined.data$thompson_PC1,combined.data$df16)
png("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/CuiRep_PC1_PFN15.png",
    width = 4000, 
    height = 4000,
    res=1000)
ggplot(data=combined.data,aes(thompson_PC1,df16))+geom_point(size=1,colour='gold')+geom_smooth(method=lm,colour='gray')+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black")) +ylab('Network 15') + xlab('Thompson PC1: General Cognition') + ggtitle(paste("r=",as.character(round(correlation$estimate,3)),", p=",as.character(round(correlation$p.value,3))))
dev.off()

correlation<-cor.test(combined.data$thompson_PC1,combined.data$df3)
png("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/CuiRep_PC1_PFN2.png",
    width = 4000, 
    height = 4000,
    res=1000)
ggplot(data=combined.data,aes(thompson_PC1,df3))+geom_point(size=1,colour='steel blue')+geom_smooth(method=lm,colour='gray')+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black")) +ylab('Network 2') + xlab('Thompson PC1: General Cognition') + ggtitle(paste("r=",as.character(round(correlation$estimate,3)),", p=",as.character(round(correlation$p.value,3))))
dev.off()

correlation<-cor.test(combined.data$interview_age,combined.data$df5)
png("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/CuiRep_Age_PFN4.png",
    width = 4000, 
    height = 4000,
    res=1000)
ggplot(data=combined.data,aes(interview_age,df5))+geom_point(size=1,colour='steel blue')+geom_smooth(method=lm,colour='gray')+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black")) +ylab('Network 4') + xlab('Age')  + ggtitle(paste("r=",as.character(round(correlation$estimate,3)),", p=",as.character(round(correlation$p.value,3))))
dev.off()

correlation<-cor.test(combined.data$interview_age,combined.data$df11)
png("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/CuiRep_Age_PFN10.png",
    width = 4000, 
    height = 4000,
    res=1000)
ggplot(data=combined.data,aes(interview_age,df11))+geom_point(size=1,colour='purple')+geom_smooth(method=lm,colour='gray')+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black")) +ylab('Network 10') + xlab('Age') + ggtitle(paste("r=",as.character(round(correlation$estimate,3)),", p=",as.character(round(correlation$p.value,3))))
dev.off()

## Attempt at predictive modeling to see if we can predict the 3 independent domains of cognition from all the PFNs
# PC1
```{r}
base_model<-lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+(1|abcd_site),data=train.data)
pfn_model<-lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+(1|abcd_site)+df2+df3+df4+df5+df6+df7+df8+df9+df10+df11+df12+df13+df14+df15+df16+df17+df18,data=train.data)
anova(base_model,pfn_model,test="Chisq")

# Lasso to just use the PFNs that are relevant:
x=model.matrix(lm(thompson_PC1~df2+df3+df4+df5+df6+df7+df8+df9+df10+df11+df12+df13+df14+df15+df16+df17+df18,data=train.data))
y=train.data[!is.na(train.data$thompson_PC1),]$thompson_PC1

#Use a second train/validation division to select the "lambda" for the lasso.
set.seed(42894)
train=sample(seq(2165),721,replace=FALSE)
lasso.tr=glmnet(x[train,],y[train])
pred=predict(lasso.tr,x[-train,])
rmse=sqrt(apply((y[-train]-pred)^2,2,mean))
plot(log(lasso.tr$lambda),rmse,type="b",xlab="Log(lambda")
lam.best=lasso.tr$lambda[order(rmse)[1]] # extracting the best lambda - this extracts that component -- order puts them in order
coef(lasso.tr,s=lam.best)

# Test whether the model significantly improves over not having PFNs included
base_model<-lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+(1|abcd_site),data=train.data)
pfn_model_lasso<-lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+(1|abcd_site)+df4+df5+df6+df7+df8+df10+df11+df13+df15+df16+df18,data=train.data)
anova(base_model,pfn_model_lasso,test="Chisq")

# Test whether the model can predict cognition in held out test data
pred.thompsonPC1<-predict(pfn_model_lasso,newdata=test.data[!is.na(test.data$thompson_PC1),c("interview_age","sex","rel_family_id","abcd_site","df4","df5","df6","df7","df8","df10","df11","df13","df15","df16","df18")],allow.new.levels=TRUE)
plot(test.data[!is.na(test.data$thompson_PC1),]$thompson_PC1,pred.thompsonPC1)
cor.test(test.data[!is.na(test.data$thompson_PC1),]$thompson_PC1,pred.thompsonPC1)
```
